---
title: "Pre-Meeting Diagnostics: Staged Sigmoidal L-MNLFA of Pubertal Development"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(knitr)
  library(kableExtra)
})

sshfs_default <- "/private/tmp/sshfs/projects/abcd-projs/dissertation/study1/outputs"
out_base <- Sys.getenv("OUT_DIR")
if (!nzchar(out_base) || !dir.exists(out_base)) {
  out_base <- if (dir.exists(sshfs_default)) sshfs_default else file.path(Sys.getenv("HOME"), "projects/abcd-projs/dissertation/study1/outputs")
}
model_dir <- file.path(out_base, "lmnlfa_growth_sigmoid_staged")
premeeting_dir <- file.path(model_dir, "premeeting")

sexes <- c("female", "male")
tag <- setNames(paste0(sexes, "_n1500"), sexes)

pth <- function(subfolder, filename_tmpl, sx) {
  file.path(premeeting_dir, subfolder, sprintf(filename_tmpl, tag[sx]))
}
read_csv_safe <- function(path) {
  if (!file.exists(path)) return(NULL)
  read.csv(path, check.names = FALSE)
}
img_tag <- function(path, alt, width = "100%") {
  if (!file.exists(path)) {
    return(sprintf('<p><em>[Figure not available: %s]</em></p>', basename(path)))
  }
  sprintf('<img src="%s" alt="%s" style="width:%s;max-width:600px;border:1px solid #ddd;padding:4px;"/>', path, alt, width)
}
kbl_html <- function(df, digits = 3, caption = NULL, highlight = FALSE) {
  if (is.null(df) || nrow(df) == 0) return('<p><em>No data.</em></p>')
  k <- kbl(df, digits = digits, caption = caption, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 12)
  if (highlight && "rhat" %in% names(df) && "ess_bulk" %in% names(df)) {
    bad <- which(df$rhat > 1.01 | df$ess_bulk < 400)
    if (length(bad) > 0) k <- row_spec(k, bad, background = "#FFF3B0")
  }
  as.character(k)
}
sex_panel <- function(label_fn) {
  cat('<div style="display:flex;gap:20px;flex-wrap:wrap;margin-bottom:24px;">\n')
  for (sx in sexes) {
    cat(sprintf('<div style="flex:1;min-width:340px;">\n<h4>%s</h4>\n', tools::toTitleCase(sx)))
    cat(label_fn(sx))
    cat('</div>\n')
  }
  cat('</div>\n')
}
missing_stage_note <- function(sx, stage_letter) {
  sprintf('<p><em>Stage %s fit not available for %s.</em></p>', stage_letter, sx)
}
questions_box <- function(qs) {
  items <- paste(sprintf('<li>%s</li>', qs), collapse = "\n")
  sprintf(
    '<div style="background:#eef6fb;border-left:4px solid #335C67;padding:8px 16px;margin:10px 0;"><strong>Questions to ask:</strong><ul style="margin-bottom:0;">%s</ul></div>',
    items
  )
}

# ---- context figures: copies of existing project figures (sample timelines,
# raw-vs-model overlays), resized (sips, macOS) so the self-contained HTML
# stays a reasonable email attachment size. Source figures are NOT modified;
# these are one-time copies written to premeeting/context/. If sips isn't
# available (non-macOS), the original (larger) file is copied as-is.
context_dir <- file.path(premeeting_dir, "context")
dir.create(context_dir, showWarnings = FALSE, recursive = TRUE)
context_sources <- c(
  timeline_modeled = file.path(out_base, "exploration", "sample_timeline", "timeline_modeled_%s.png"),
  raw_vs_model = file.path(out_base, "viz_raw_vs_model", "raw_vs_model_%s.png"),
  raw_traj_youth = file.path(out_base, "viz_raw_vs_model", "raw_trajectories_modeled_sample_%s_youth.png"),
  raw_traj_caregiver = file.path(out_base, "viz_raw_vs_model", "raw_trajectories_modeled_sample_%s_caregiver.png")
)
for (nm in names(context_sources)) {
  for (sx in sexes) {
    src <- sprintf(context_sources[[nm]], sx)
    dest <- file.path(context_dir, basename(src))
    if (file.exists(src) && (!file.exists(dest) || file.mtime(src) > file.mtime(dest))) {
      file.copy(src, dest, overwrite = TRUE)
      if (nzchar(Sys.which("sips"))) system2("sips", c("-Z", "1200", shQuote(dest)), stdout = FALSE, stderr = FALSE)
    }
  }
}
ctx_path <- function(nm, sx) file.path(context_dir, basename(sprintf(context_sources[[nm]], sx)))

# ---- load everything needed for verdicts up front ----
waterfall <- lapply(sexes, function(sx) read_csv_safe(pth("data-audit", "waterfall_%s.csv", sx)))
names(waterfall) <- sexes
persons_per_wave <- lapply(sexes, function(sx) read_csv_safe(pth("data-audit", "persons_per_wave_%s.csv", sx)))
names(persons_per_wave) <- sexes

wave_dip_note <- sapply(sexes, function(sx) {
  d <- persons_per_wave[[sx]]
  if (is.null(d)) return(sprintf("%s: wave data not available.", sx))
  d2 <- d %>% group_by(wave) %>% summarize(n_persons = sum(n_persons), .groups = "drop") %>% filter(wave != "bl")
  dip <- d2[which.min(d2$n_persons), ]
  others <- d2$n_persons[d2$wave != dip$wave]
  sprintf("%s: %s is the low point (n=%d combined caregiver+youth) vs %d-%d across the other follow-up waves", tools::toTitleCase(sx), dip$wave, dip$n_persons, min(others), max(others))
})
race_counts <- lapply(sexes, function(sx) read_csv_safe(pth("data-audit", "race_group_counts_%s.csv", sx)))
names(race_counts) <- sexes
baseline_wave <- lapply(sexes, function(sx) read_csv_safe(pth("data-audit", "baseline_wave_origin_%s.csv", sx)))
names(baseline_wave) <- sexes

ceil_floor <- lapply(sexes, function(sx) read_csv_safe(pth("ceiling-floor", "ceiling_floor_by_item_%s.csv", sx)))
names(ceil_floor) <- sexes
all_ceil <- lapply(sexes, function(sx) read_csv_safe(pth("ceiling-floor", "all_items_ceiling_floor_%s.csv", sx)))
names(all_ceil) <- sexes
infl_summ <- lapply(sexes, function(sx) read_csv_safe(pth("ceiling-floor", "inflection_age_summary_%s.csv", sx)))
names(infl_summ) <- sexes
pop_curve <- lapply(sexes, function(sx) read_csv_safe(pth("ceiling-floor", "population_mean_curve_range_%s.csv", sx)))
names(pop_curve) <- sexes

conv_all <- lapply(sexes, function(sx) read_csv_safe(pth("convergence", "convergence_all_params_%s.csv", sx)))
names(conv_all) <- sexes
flagged <- lapply(sexes, function(sx) read_csv_safe(pth("convergence", "flagged_params_%s.csv", sx)))
names(flagged) <- sexes
diagtab <- lapply(sexes, function(sx) read_csv_safe(pth("convergence", "divergences_treedepth_%s.csv", sx)))
names(diagtab) <- sexes
bcor <- lapply(sexes, function(sx) read_csv_safe(pth("convergence", "stageB_growth_age_dif_correlations_%s.csv", sx)))
names(bcor) <- sexes

sens_grid <- lapply(sexes, function(sx) read_csv_safe(pth("dif-sensitivity", "dif_sensitivity_grid_summary_%s.csv", sx)))
names(sens_grid) <- sexes
raw_dif <- lapply(sexes, function(sx) read_csv_safe(pth("dif-sensitivity", "stageB_raw_dif_posteriors_%s.csv", sx)))
names(raw_dif) <- sexes

impact_cov <- lapply(sexes, function(sx) read_csv_safe(pth("impact-comparison", "stageA_vs_C_impact_covariates_%s.csv", sx)))
names(impact_cov) <- sexes
impact_scal <- lapply(sexes, function(sx) read_csv_safe(pth("impact-comparison", "stageA_vs_C_impact_scalars_%s.csv", sx)))
names(impact_scal) <- sexes

agree <- lapply(sexes, function(sx) read_csv_safe(pth("agreement", "parent_youth_agreement_%s.csv", sx)))
names(agree) <- sexes

# menarche (fpete) is female-only; these files only exist for sx == "female"
item5_completeness <- read_csv_safe(pth("item5-exploration", "item5_completeness_vs_other_items_%s.csv", "female"))
item5_missing <- read_csv_safe(pth("item5-exploration", "item5_missingness_cost_%s.csv", "female"))
item5_rate <- read_csv_safe(pth("item5-exploration", "item5_rate_by_age_%s.csv", "female"))
item5_direction <- read_csv_safe(pth("item5-exploration", "item5_direction_summary_%s.csv", "female"))
item5_steepness <- read_csv_safe(pth("item5-exploration", "item5_steepness_comparison_%s.csv", "female"))

verdict_item5 <- if (is.null(item5_direction)) {
  "Menarche diagnostics not available for this render."
} else {
  sprintf(
    "Menarche's raw coding runs %s with age (%s the usual PDS 'higher code = more developed' convention), which %s explain the earlier Stage A/B/C convergence failure on its own; it is also a much steeper (near-step-function) item than the four already modeled. Neither issue requires a Stan-file change to fix -- only a coding correction, a possibly wider loading prior, and re-adding it to build_lmnlfa_data_staged().",
    item5_direction$direction[1],
    if (item5_direction$coding_matches_assumption[1]) "matching" else "OPPOSITE of",
    if (!item5_direction$coding_matches_assumption[1]) "plausibly does" else "does not"
  )
}

# ---- verdict computations (numbers only; sentence framing is fixed logic) ----
max_ceiling_any <- max(unlist(lapply(all_ceil, function(x) if (!is.null(x)) max(x$prop_all_ceiling) else NA)), na.rm = TRUE)
verdict_ceiling <- if (max_ceiling_any > 0.5) {
  sprintf("A fixed common ceiling looks reasonable for the oldest respondents -- up to %.0f%% of person-occasions have all four items at the top category in the oldest/most-developed cells.", 100 * max_ceiling_any)
} else {
  sprintf("A fixed common ceiling is NOT clearly reached for most of the sample -- even the oldest-age, most-developed cell tops out at only %.0f%% of person-occasions with all four items at the top category.", 100 * max_ceiling_any)
}

max_growth_dif_cor <- max(unlist(lapply(bcor, function(x) {
  if (is.null(x)) return(NA)
  rows <- x[x$param %in% c("mu_logk", "mu_alpha"), ]
  cols <- setdiff(names(x), c("param", "mu_logk", "mu_alpha"))
  max(abs(as.matrix(rows[, cols])))
})), na.rm = TRUE)
verdict_identifiability <- if (max_growth_dif_cor > 0.5) {
  sprintf("Stage B's growth-curve parameters and age DIF are NOT cleanly separated -- the largest posterior correlation between a growth mean parameter and an age DIF term reaches %.2f across the two sexes.", max_growth_dif_cor)
} else {
  sprintf("Stage B's growth-curve parameters and age DIF appear reasonably well identified -- the largest posterior correlation between a growth mean parameter and an age DIF term is %.2f across the two sexes.", max_growth_dif_cor)
}

sens_range_ratio <- max(unlist(lapply(sens_grid, function(x) {
  if (is.null(x)) return(NA)
  rng <- range(x$n_loading_retained)
  if (rng[2] == 0) return(0)
  (rng[2] - rng[1]) / rng[2]
})), na.rm = TRUE)
verdict_dif_robust <- if (sens_range_ratio > 0.5) {
  sprintf("The retained DIF set is sensitive to screening thresholds -- the number of retained loading-DIF terms swings by more than %.0f%% of its maximum across the grid of FDR/magnitude-floor settings tested.", 100 * sens_range_ratio)
} else {
  sprintf("The retained DIF set is reasonably robust to screening thresholds -- the number of retained loading-DIF terms varies by at most %.0f%% of its maximum across the grid of FDR/magnitude-floor settings tested.", 100 * sens_range_ratio)
}

max_impact_shift <- max(unlist(lapply(impact_cov, function(x) if (!is.null(x)) max(abs(x$diff_as_frac_A_sd)) else NA)), na.rm = TRUE)
verdict_impact_bias <- if (max_impact_shift > 1) {
  sprintf("Stage A impact estimates ARE materially affected by ignoring DIF -- the largest Stage A-to-C shift reaches %.1f Stage-A posterior SDs.", max_impact_shift)
} else {
  sprintf("Stage A impact estimates are NOT materially biased by ignoring DIF -- the largest Stage A-to-C shift across both sexes is %.1f Stage-A posterior SDs.", max_impact_shift)
}

mean_spearman <- mean(unlist(lapply(agree, function(x) if (!is.null(x)) mean(x$spearman_r, na.rm = TRUE) else NA)), na.rm = TRUE)
mean_within1 <- mean(unlist(lapply(agree, function(x) if (!is.null(x)) weighted.mean(x$within_one, x$n) else NA)), na.rm = TRUE)
verdict_agreement <- if (mean_spearman > 0.3 && mean_within1 > 0.85) {
  sprintf("Parent and youth reports plausibly track the same latent quantity -- mean within-one-category agreement is %.0f%% and mean Spearman correlation is %.2f across items/age bands/sexes.", 100 * mean_within1, mean_spearman)
} else {
  sprintf("Parent-youth agreement is modest enough to question a shared latent quantity assumption without informant DIF adjustment -- mean within-one-category agreement is %.0f%% and mean Spearman correlation is only %.2f across items/age bands/sexes.", 100 * mean_within1, mean_spearman)
}
```

## Context

Reporting on my staged, sigmoidal longitudinal moderated nonlinear factor analysis (L-MNLFA) of pubertal development in the ABCD Study, using the Pubertal Development Scale as reported by both caregivers and youth. Puberty is modeled as a bounded logistic growth curve with a fixed floor (1) and ceiling (5), and two correlated person-specific random effects -- a log-rate ("tempo") and an inflection age ("timing") -- estimated in three stages: growth-factor impact only (Stage A), DIF screening with impact fixed (Stage B), and a final combined fit with impact free and only the screened DIF terms retained (Stage C). Caregiver and youth reports enter as separate informant-level rows, so reporter-related measurement bias can be tested directly as DIF alongside age, race/ethnicity, and waist-to-height ratio (WtHR). Female and male results are shown side by side in every section below.

---

## 0. Sample & raw-data context

**Verdict:** The raw (unmodeled) score trajectories visibly flatten with age in a way broadly consistent with a sigmoid, and the n=1,500 modeled subsample's wave-coverage pattern mirrors the full sample's -- including the same fu3 dip, which reflects a real assessment-timing gap in the data, not an artifact of subsampling.

```{r sec0, results='asis'}
sex_panel(function(sx) {
  out <- paste0(
    "<p><strong>Sample timeline (n=1,500 modeled subsample)</strong> -- each row is one person, colored by wave completed</p>",
    img_tag(ctx_path("timeline_modeled", sx), paste("Sample timeline,", sx))
  )
  out <- paste0(out, "<p><strong>Raw vs. fitted model trajectory</strong></p>", img_tag(ctx_path("raw_vs_model", sx), paste("Raw vs model,", sx)))
  out <- paste0(out, "<p><strong>Raw item trajectories, modeled sample -- youth-reported</strong></p>", img_tag(ctx_path("raw_traj_youth", sx), paste("Raw trajectories youth,", sx)))
  out <- paste0(out, "<p><strong>Raw item trajectories, modeled sample -- caregiver-reported</strong></p>", img_tag(ctx_path("raw_traj_caregiver", sx), paste("Raw trajectories caregiver,", sx)))
  out
})
cat(questions_box(c(
  "Does the raw (unmodeled) trajectory shape in <code>raw_vs_model_&lt;sex&gt;.png</code> look sigmoidal to you -- i.e. does the model's functional-form choice look right on the raw data, independent of the fitted curve?",
  sprintf("Wave coverage dips at a specific follow-up wave in both sexes (%s; %s) -- does it warrant a robustness check excluding or reweighting that wave?", wave_dip_note[1], wave_dip_note[2]),
  "Do the youth- vs. caregiver-reported raw trajectories look like they're tracking the same underlying process?"
)))
```

---

## 1. Data audit

**Verdict:** Roughly two-thirds to three-quarters of raw person-occasion rows survive the full cleaning pipeline (age/race/WtHR completeness and range, item completeness, mapped race group), and a small (<5%) share of the analysis sample's "baseline" WtHR value actually comes from a later wave because true baseline was filtered out for those people.

```{r sec1, results='asis'}
sex_panel(function(sx) {
  out <- ""
  wf <- waterfall[[sx]]
  if (!is.null(wf)) out <- paste0(out, "<p><strong>Filtering waterfall</strong></p>", kbl_html(wf, digits = 0))
  bw <- baseline_wave[[sx]]
  if (!is.null(bw)) {
    n_not_bl <- sum(bw$n_persons[!bw$is_baseline_wave])
    pct <- round(100 * n_not_bl / sum(bw$n_persons), 1)
    out <- paste0(out, sprintf("<p><strong>Baseline-occasion wave origin</strong> (%d people, %.1f%%, drawn from a non-bl wave)</p>", n_not_bl, pct), kbl_html(bw, digits = 0))
  }
  rc <- race_counts[[sx]]
  if (!is.null(rc)) out <- paste0(out, "<p><strong>Race group counts (analysis sample)</strong></p>", kbl_html(rc, digits = 0))
  out
})
cat(questions_box(c(
  "More of a note to myself: Roughly a third of raw rows are dropped in the filtering waterfall above -- we should check whether dropped rows are systematically different by race or site, given Hispanic and Black are already the two smallest retained race groups.",
  "Is the share of people whose 'baseline' WtHR is drawn from a non-bl wave (see the baseline-occasion table above) small enough to ignore, or should those people be flagged/excluded in a sensitivity check on the WtHR impact estimate?"
)))
```

---

## 2. Ceiling / floor support

**Verdict:** `r verdict_ceiling`

```{r sec2, results='asis'}
sex_panel(function(sx) {
  out <- ""
  pc <- pop_curve[[sx]]
  if (!is.null(pc)) {
    out <- paste0(out, "<p><strong>Population-mean curve at observed age extremes</strong></p>", kbl_html(pc, digits = 3))
  }
  isumm <- infl_summ[[sx]]
  if (!is.null(isumm)) {
    out <- paste0(out, "<p><strong>Inflection-age distribution</strong></p>", kbl_html(isumm, digits = 2))
    out <- paste0(out, img_tag(pth("ceiling-floor", "inflection_age_histogram_%s.png", sx), paste("Inflection age histogram,", sx)))
  }
  out <- paste0(out, "<p><strong>All-items-at-ceiling/floor by age band</strong></p>")
  ac <- all_ceil[[sx]]
  if (!is.null(ac)) out <- paste0(out, kbl_html(ac, digits = 3))
  if (sx == "male") {
    out <- paste0(out, '<p><em>Facial hair (mpete) is included as a 5th item below, matched to the same modeled person-occasions -- it is decided to be added to the full model on the next rerun.</em></p>')
  }
  out <- paste0(out, img_tag(pth("ceiling-floor", "ceiling_floor_by_item_%s.png", sx), paste("Ceiling/floor by item,", sx)))
  out
})
cat(questions_box(c(
  "Male item-level ceiling rates at 15+ are consistently lower than female (see the ceiling/floor-by-item figure above and Section 0's raw trajectories) -- reflects a real sex difference in development, and/or is ABCD's age range simply too truncated to observe most males finishing puberty, which would argue for caution interpreting phi_alpha (inflection-age variance) for males?",
  "None of the individual inflection ages fall outside the observed age range for either sex -- given that, is a FIXED floor/ceiling still preferred over freeing the ceiling as an estimated parameter, now that we can see it isn't forcing extrapolation?"
)))
```

---

## 3. Convergence, focused on the DIF-vs-growth confound

**Verdict:** `r verdict_identifiability`

```{r sec3, results='asis'}
sex_panel(function(sx) {
  out <- ""
  dt <- diagtab[[sx]]
  if (!is.null(dt)) out <- paste0(out, "<p><strong>Divergences / max-treedepth by stage</strong></p>", kbl_html(dt, digits = 0))
  fl <- flagged[[sx]]
  if (!is.null(fl)) {
    out <- paste0(out, sprintf("<p><strong>%d flagged parameters</strong> (rhat &gt; 1.01 or ess_bulk &lt; 400)</p>", nrow(fl)))
    out <- paste0(out, kbl_html(head(fl, 20), digits = 3, highlight = TRUE))
    if (nrow(fl) > 20) out <- paste0(out, sprintf("<p><em>(%d more in the appendix)</em></p>", nrow(fl) - 20))
  }
  out <- paste0(out, "<p><strong>Stage B: growth params vs age-DIF posterior correlations</strong></p>")
  bc <- bcor[[sx]]
  if (!is.null(bc)) out <- paste0(out, kbl_html(bc, digits = 2))
  out <- paste0(out, img_tag(pth("convergence", "stageB_growth_age_dif_cor_heatmap_%s.png", sx), paste("Growth x age-DIF correlation heatmap,", sx)))
  out
})
cat(questions_box(c(
  "The mu_alpha/age-loading-DIF correlations in the heatmap above reach the mid-.80s for males -- is BH-FDR screening on Stage B adequate given that level of confounding, or does this specific finding argue for running the already-built spike-and-slab alternative (joint shrinkage instead of post-hoc screening) before trusting Stage C's retained age-DIF pattern?",
  "Since age is both the growth curve's own axis and a DIF covariate, would you rather see age-varying item loadings modeled as part of the growth structure itself (e.g., a smooth loading-by-age term) rather than as a separate DIF effect competing with mu_alpha for the same variation?"
)))
```

<details><summary><strong>Appendix: full per-stage parameter tables</strong></summary>

```{r sec3-appendix, results='asis'}
for (sx in sexes) {
  cat(sprintf("<h4>%s</h4>\n", tools::toTitleCase(sx)))
  ca <- conv_all[[sx]]
  if (!is.null(ca)) cat(kbl_html(ca, digits = 3, highlight = TRUE))
}
```

</details>

---

## 4. DIF screening sensitivity

**Verdict:** `r verdict_dif_robust`

```{r sec4, results='asis'}
sex_panel(function(sx) {
  out <- "<p><strong>Retained DIF terms across the FDR-q x magnitude-floor grid</strong></p>"
  sg <- sens_grid[[sx]]
  if (!is.null(sg)) out <- paste0(out, kbl_html(sg, digits = 3))
  out <- paste0(out, img_tag(pth("dif-sensitivity", "dif_sensitivity_heatmap_%s.png", sx), paste("DIF sensitivity heatmap,", sx)))
  out
})
cat(questions_box(c(
  "The male grid (see the sensitivity table above) drops from 10 to 0 retained loading-DIF terms between a magnitude floor of .05 and .2 -- independent of what maximizes retained terms, what magnitude floor would you consider defensible for this response scale?",
  "Given the retained set changes this much across plausible threshold choices, should the write-up report a single 'selected' DIF pattern, or present Stage C results under 2-3 alternative screening thresholds as a robustness check?"
)))
```

<details><summary><strong>Appendix: raw Stage B DIF posterior means and 90% CIs</strong></summary>

```{r sec4-appendix, results='asis'}
for (sx in sexes) {
  cat(sprintf("<h4>%s</h4>\n", tools::toTitleCase(sx)))
  rd <- raw_dif[[sx]]
  if (!is.null(rd)) cat(kbl_html(rd, digits = 3))
}
```

</details>

---

## 5. Stage A vs Stage C impact

**Verdict:** `r verdict_impact_bias`

```{r sec5, results='asis'}
sex_panel(function(sx) {
  out <- "<p><strong>By covariate (growth-curve impact)</strong></p>"
  ic <- impact_cov[[sx]]
  if (!is.null(ic)) out <- paste0(out, kbl_html(ic, digits = 3))
  out <- paste0(out, "<p><strong>Scalar growth parameters</strong></p>")
  isc <- impact_scal[[sx]]
  if (!is.null(isc)) out <- paste0(out, kbl_html(isc, digits = 3))
  out <- paste0(out, img_tag(pth("impact-comparison", "stageA_vs_C_impact_plot_%s.png", sx), paste("Stage A vs C impact,", sx)))
  out
})
```

---

## 6. Caregiver-youth agreement

**Verdict:** `r verdict_agreement`

```{r sec6, results='asis'}
sex_panel(function(sx) {
  out <- "<p><strong>Agreement by item x age band</strong></p>"
  ag <- agree[[sx]]
  if (!is.null(ag)) out <- paste0(out, kbl_html(ag, digits = 3))
  out <- paste0(out, img_tag(pth("agreement", "parent_youth_mean_diff_%s.png", sx), paste("Mean signed caregiver-youth difference,", sx)))
  out
})
cat(questions_box(c(
  "Exact agreement is only ~56% while within-one-category agreement is ~93% -- should the informant DIF terms already retained in Stage C be interpreted as correcting genuine reporter bias, or could collapsing PDS items to fewer response categories be a simpler, more defensible fix?",
  "Does the mean-signed-difference plot above show a trend across age bands (reporter divergence changing over development), or is it flat enough to treat informant bias as roughly constant with age?"
)))
```

---

## 7. Menarche (fpete): diagnosing the convergence failure

Menarche (`fpete`, binary, female only) previously caused Stage A/B/C convergence problems when included, so it isn't in the current model. The diagnostics below exist to diagnose *why* menarche broke convergence.

**Verdict:** `r verdict_item5`

```{r sec7, results='asis'}
out <- ""
if (!is.null(item5_direction)) {
  flag <- if (!item5_direction$coding_matches_assumption[1]) {
    '<p style="color:#9E2A2B;"><strong>Coding direction is INVERTED</strong> relative to the other PDS items -- code %d (not the larger number) is the "developed/reached" response.</p>'
  } else {
    '<p>Coding direction matches the other PDS items (higher code = more developed).</p>'
  }
  out <- paste0(out, "<p><strong>Candidate cause 1 -- coding direction:</strong> raw response-rate trend with age is '", item5_direction$direction[1], "'.</p>",
    sprintf(flag, item5_direction$developed_code[1]))
}
if (!is.null(item5_steepness)) {
  item5_jump <- max(abs(diff(item5_rate$prop_developed[item5_rate$reporter == "parent"])))
  out <- paste0(out, sprintf(
    "<p><strong>Candidate cause 2 -- steepness:</strong> menarche's largest single-age-band jump in 'developed' proportion is %.2f, vs. an average of %.2f across the four modeled items' ceiling proportions (~%.1fx steeper).</p>",
    item5_jump, mean(item5_steepness$max_jump), item5_jump / mean(item5_steepness$max_jump)
  ))
  out <- paste0(out, "<p><em>Steepness comparison (largest age-band jump, by item)</em></p>", kbl_html(
    bind_rows(item5_steepness, tibble(item = "Menarche (fpete)", max_jump = item5_jump)), digits = 3
  ))
}
ct <- item5_completeness
if (!is.null(ct)) out <- paste0(out, "<p><strong>Completeness vs. the four modeled items</strong> (among age/race/WtHR-filtered rows) -- ruled out as a contributing cause</p>", kbl_html(ct, digits = 1))
mt <- item5_missing
if (!is.null(mt)) out <- paste0(out, "<p><em>Cost of also requiring menarche</em> (among occasions already passing the current 4-item filter)</p>", kbl_html(mt, digits = 1))
rt <- item5_rate
if (!is.null(rt)) out <- paste0(out, "<p><strong>Response rate by age band x informant</strong></p>", kbl_html(rt, digits = 3))
out <- paste0(out, img_tag(pth("item5-exploration", "item5_rate_by_age_%s.png", "female"), "Menarche rate by age"))
cat(sprintf('<div style="max-width:700px;">%s</div>', out))
cat(questions_box(c(
  "Does the coding-direction inversion alone plausibly account for the earlier non-convergence, or is the steepness (near-step-function transition) likely a separate, additional contributor that would need its own fix (e.g. a wider loading prior) even after the coding is corrected?"
)))
```

---

## Open decisions

- **Fixed common ceiling.** `r verdict_ceiling`
- **Stage B identifiability of growth curve vs age DIF.** `r verdict_identifiability`
- **Robustness of the retained DIF set.** `r verdict_dif_robust`
- **Shared latent eta across informants.** `r verdict_agreement`
